Consistency scores in text data

### 2020-08-10 outputs/documents/paper.Rmd
brackets correction

To illustrate the difference, consider the quote from Descartes, 'I think therefore I am'. The focus word for which we want to learn the embedding is 'therefore', and the context window is set to two. This means that we consider the two words immediately preceding and proceeding 'therefore'. The CBOW input is 'I, think, I, am', and the output is 'therefore'. In contrast, the continuous skip-gram input is 'therefore', while 'I, think, I, am' is the output. The input layer encodes terms within a context window to one-hot vectors. Using a projection matrix, the projection layer then project the discrete one-hot vectors to lower-dimension continuous vectors. Finally, the output layer is a softmax regression classifier that obtains the posterior distribution of words (@rong2014word2vec). It is the input and output vectors that differ between the two architectures.

Word2Vec is trained on the Google News dataset (@citegooglenews). This contains six billion tokens and the vocabulary size is restricted to the one million most common words (@mikolov2013efficient). As this is a large dataset, it provides an accurate statistical pattern of language, as it is used in this dataset, and a comprehensive mapping between terms and vector positions.

@@

There are a few packages for implementing a Word2Vec model in R. The packages are `rword2vec` by @rword2vec and `word2vec` by @word2vecR. Both packages use the same algorithm that Word2Vec uses to produce word representations, but they do not have an affiliation with Google. Another way to implement a Word2Vec model in R is through Keras package. Keras is an API for deep learning written in Python (@chollet2015keras) and the Keras package provides an R interface to Keras (@kerasR2020). RStudio published a tutorial for implementing the skip-gram model of Word2Vec (@falbel2017word) through Keras. We intend to follow this tutorial as the approach to implement a Word2Vec model. There are three basic steps within the Keras deep learning framework: first is to prepare the data, second is to define the model and third is to train and evaluate the model. Following the steps, we first downloaded a the [Amazon Fine Foods Reviews dataset from Stanford] (@amzfinefooddata) and preprocess the text into tokens. Then we defined a model that mimics the skip-gram algorithm of Word2Vec in a function called `skipgrams_generator()`. The embedding dimensions, the size of the context window and the number of samples are configurable within this function. Finally, we used the `fit_generator()` function of Keras to train the model. **What's supposed to happen:** As a result, a embedding matrix is obtained and serves as the look-up table of the word vectors. **What really happened:** However, the training process was stuck in the operation of `fit_generator()` function. According to Stack Flow this might be an issue of incompatible versions between Keras and Tensorflow. More investigation is needed.

@@

There are more than one way to implement GloVe word embedding model in R. One way is to load GloVe embeddings into a frozen Keras Embedding layer (@kerasglove) using the `Keras` R package (@kerasR2020). Another way, as mentioned and briefly demonstrated by Rstudio in their blog post, is to use the `text2vec` package developed by @text2vec. Keras published a blog post on pre-trained word embeddings through GloVe in R. In this project, we intend to follow the blog post for implementing GloVe through the Keras R package (@kerasglove). The blog post is aimed to create a word embedding model for text classification. However, we will only use the guide create a embedding index and skip the steps of text classification. Instead of building a model for word embedding from scratch as we did for the skip-gram model of Word2Vec, here we are merely loading the Glove model and its pre-trained word vectors as an index. The word embedding index is then saved to an R environment.

@@

OpenAI is an AI research company who openly pursues Artificial General Intelligence — to have machines learn and understand any intellectual tasks as human do (@brundage2018malicious). OpenAI's Generative Pre-Trained Transformer (GPT) models are pre-trained language models that are built based on the transformer architecture. Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based and CNN-based models in computational efficiency (@vaswani2017attention). Since the emergence of the Transformer model, most of the representative pre-trained models are built based on this architecture (@hanretty2018comparing). Therefore, we can say that Transformer marks a new era of language models.

As of August 2020, OpenAI has three generations of GPT models. GPT was introduced to the public in June, 2018. In the year 2019, its successor GPT-2 was released in three stages. Not long after, in June 2020, OpenAI announced GPT-3, which is the latest GPT model as of July 2020 (@brown2020language). Since the API of GPT-3 at the moment is only available to a few invited users, public applications of GPT-3 are limited. On the other hand, its predecessor, GPT-2, has enabled researchers made breakthroughs in the past years. @alt2019fine extended GPT-2 to relation extraction, which is a task of identifying the relationship between concepts appeared in text. They had shown a milestone in predicting larger range of distinct relation types with higher confidence (@alt2019fine). In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning (@bird2020lstm). @bird2020lstm employed GPT-2 as part of the solution to generate augmented data for classification. GPT-2 also helped NLP researchers tackling the task of textual paraphrasing. @hegde2020unsupervised used GPT-2 to build an unsupervised paraphrasing model that is not restricted by domain shift within an article, which is a problem faced by the usual supervised models of paraphrasing.

Most of prior language models are trained in a single domain by supervised learning to perform a single task. OpenAI, on the other hands, train the GPT models to perform many tasks (@radford2019language). GPT-2 is pre-trained on WebText, a comprehensive dataset that is scrapped from millions of webpages and contains total of 40 GB of text (@radford2019language). WebText contains all kinds of textual data: dialogues, comments, translations, questions and answers...etc. Such diversity is critical for GPT-2 to be a generalized model that can perform multiple tasks with any domain boundary. However, GPT-2 has the capacity to adapt to a context and generates custom output with the "fine-tuning" technique (@ziegler2019fine). While the pre-training stage is unsupervised and mandatory, fine-tuning is an optional training stage where GPT can learn to perform new tasks in a supervised or reinforced approach such as stylistic continuation and summarization (@ziegler2019fine). During the pre-training stage, the model is trained to perform text completion task. Give it a prompt to start and it will generate texts that continues what has been started. In the fine-tuning stage, @radford2019language provided additional training datasets for GPT-2 to generate the texts with specific sentiment, or be descriptive, or even summarize the prompt text.

OpenAI has made GPT-2 open-sourced through Github with Python and Tensorflow (@tensorflow2015-whitepaper). There is also a package in R that wraps the Python code into a R package, created by @keydanaluraschi2019gpt2. There are three choices of models in terms of parameter sizes — "124M", "355M" and "774M". The R package only permits exploratory experiments with the pre-trained tasks; fine-tuning for new tasks is not available thorough R. To proceed with the implementation of GPT-2 in R, theoretically, we will fine-tune the model through Python and create a wrapper code for R. This could possibility be achieved by employing the `reticulate` R package (@reticulateR) that allows operation of Python code in R. The practicality of this approach is to be further investigated.

GPT-3 has over 175 billions of parameters (@brown2020language). In the past few years, the growth of parameters of transformer language models expands rapidly. The first generation GPT has 110 million parameters. Released in 2018, GPT-2 has 1.6 billion parameters, and within two years, GPT-3 has grown the size to 175 billion, which is over 100 times than its predecessor. The more parameters there are the better the model can generalize to new tasks; therefore, the increasing amount of parameters accelerates the improvements in text synthesis and downstream NLP tasks (@brown2020language). GPT-3 can learn NLP tasks with just a handful of example; OpenAI refers this ability as few-shot learning, or in-context learning (@brown2020language). As a result, we only need to provide GPT-3 a few demonstrations for it to learn a new NLP task. For example, OpenAI team trained GPT-3 to generate "news articles". The prompt input is a plausible first sentence of a news story written by human, and the output is an entire news article generated by the model itself. Without the demonstrations, GPT-3 tends to treat the prompts as "tweets" and generate texts that serve as responses or follow-up tweets. The team fed a few previous news articles in the model’s context as demonstrations, and the model learned to produce the outputs that adapt to the style and length of news stories (@brown2020language). This few-shot learning ability indicates that the model can learn like human do and would democratize the use of language models to broader fields.

@@

BERT stands for Bidirectional Encoder Representations from Transformers (@devlin2018bert). It was released by researchers at Google AI Language in 2018. The key innovation of BERT is its application of bidirectional training of the Transformer model, which results in a deeper sense of language context and flow. What is bidirectional training? In the context of English language, a sentence is constructed from left to right. The context on the left is used to determine the next word on the right. In this sense, GPT models are unidirectional since the probability of the next word is dependent only on the past. It is also possible to reverse the direction and model the sentence from right to left, and BERT incorporates both directions of left-to-right and right-to-left to model bidirectional contexts.

BERT is pre-trained by using two unsupervised tasks. The first task is Masked Language Model and and second task is Next Sentence Prediction. The bidirectional approach of BERT is alleviated by the "Masked Language Model" (@devlin2018bert). The Masked Language Model masks certain words in the input corpus which forces the model to predict the original words by "filling in the blank". This training process blends the representation from the left and right context and therefore increases the model's context sensitivity (@nozza2020mask). In the Next Sentence Prediction process, the model receives two sentences in sequence as input. 50% of the sentences are paired in the right sequence and the ones from the other half have sentences that are paired randomly (@devlin2018bert). The goal of this process is to have BERT evaluate if the second sentence is the next sentence of the first sentence in the original text and sharpen its capacity to predict the likelihood of the next sentence.

BERT was pre-trained on BooksCorpus dataset with 800 million words and English Wikipedia with 2,500 million words (@devlin2018bert). Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available with one additional output layer and a task-specific input (@devlin2018bert). Google AI researchers have fine-tuned BERT for 11 NLP tasks including question-answering and common-sense inference.

BERT implementation in R is also possible through Keras. @kerasbert created a python module called `keras-bert` and the module can be called by R through `reticulate` package (@reticulateR), which enables the operation of Python modules in R. To start, we install the keras-bert model in Python on the machine and load BERT model into R through `reticulate`. Once `keras-bert` is loaded, we can use it to fine-tune BERT on a task-specific training dataset. Fine-tuning is started by tokenizing the training data. The output of the tokenization is then fed to BERT as the input. The next step is to define the model by configuring its parameters such as batch size, number of epochs and the learning rate. Finally, the model is compiled and we can begin the fine-tuning.

>>>>

To illustrate the difference, consider the quote from Descartes, 'I think therefore I am'. The focus word for which we want to learn the embedding is 'therefore', and the context window is set to two. This means that we consider the two words immediately preceding and proceeding 'therefore'. The CBOW input is 'I, think, I, am', and the output is 'therefore'. In contrast, the continuous skip-gram input is 'therefore', while 'I, think, I, am' is the output. The input layer encodes terms within a context window to one-hot vectors. Using a projection matrix, the projection layer then project the discrete one-hot vectors to lower-dimension continuous vectors. Finally, the output layer is a softmax regression classifier that obtains the posterior distribution of words [@rong2014word2vec]. It is the input and output vectors that differ between the two architectures.

Word2Vec is trained on the Google News dataset [@citegooglenews]. This contains six billion tokens and the vocabulary size is restricted to the one million most common words [@mikolov2013efficient]. As this is a large dataset, it provides an accurate statistical pattern of language, as it is used in this dataset, and a comprehensive mapping between terms and vector positions.

@@

There are a few packages for implementing a Word2Vec model in R. The packages are `rword2vec` by @rword2vec and `word2vec` by @word2vecR. Both packages use the same algorithm that Word2Vec uses to produce word representations, but they do not have an affiliation with Google. Another way to implement a Word2Vec model in R is through Keras package. Keras is an API for deep learning written in Python [@chollet2015keras] and the Keras package provides an R interface to Keras [@kerasR2020]. RStudio published a tutorial for implementing the skip-gram model of Word2Vec [@falbel2017word] through Keras. We intend to follow this tutorial as the approach to implement a Word2Vec model. There are three basic steps within the Keras deep learning framework: first is to prepare the data, second is to define the model and third is to train and evaluate the model. Following the steps, we first downloaded a the [Amazon Fine Foods Reviews dataset from Stanford] [@amzfinefooddata] and preprocess the text into tokens. Then we defined a model that mimics the skip-gram algorithm of Word2Vec in a function called `skipgrams_generator()`. The embedding dimensions, the size of the context window and the number of samples are configurable within this function. Finally, we used the `fit_generator()` function of Keras to train the model. **What's supposed to happen:** As a result, a embedding matrix is obtained and serves as the look-up table of the word vectors. **What really happened:** However, the training process was stuck in the operation of `fit_generator()` function. According to Stack Flow this might be an issue of incompatible versions between Keras and Tensorflow. More investigation is needed.

@@

There are more than one way to implement GloVe word embedding model in R. One way is to load GloVe embeddings into a frozen Keras Embedding layer [@kerasglove] using the `Keras` R package [@kerasR2020]. Another way, as mentioned and briefly demonstrated by Rstudio in their blog post, is to use the `text2vec` package developed by @text2vec. Keras published a blog post on pre-trained word embeddings through GloVe in R. In this project, we intend to follow the blog post for implementing GloVe through the Keras R package [@kerasglove]. The blog post is aimed to create a word embedding model for text classification. However, we will only use the guide create a embedding index and skip the steps of text classification. Instead of building a model for word embedding from scratch as we did for the skip-gram model of Word2Vec, here we are merely loading the Glove model and its pre-trained word vectors as an index. The word embedding index is then saved to an R environment.

@@

OpenAI is an AI research company who openly pursues Artificial General Intelligence — to have machines learn and understand any intellectual tasks as human do [@brundage2018malicious]. OpenAI's Generative Pre-Trained Transformer (GPT) models are pre-trained language models that are built based on the transformer architecture. Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based and CNN-based models in computational efficiency [@vaswani2017attention]. Since the emergence of the Transformer model, most of the representative pre-trained models are built based on this architecture [@hanretty2018comparing]. Therefore, we can say that Transformer marks a new era of language models.

As of August 2020, OpenAI has three generations of GPT models. GPT was introduced to the public in June, 2018. In the year 2019, its successor GPT-2 was released in three stages. Not long after, in June 2020, OpenAI announced GPT-3, which is the latest GPT model as of July 2020 [@brown2020language]. Since the API of GPT-3 at the moment is only available to a few invited users, public applications of GPT-3 are limited. On the other hand, its predecessor, GPT-2, has enabled researchers made breakthroughs in the past years. @alt2019fine extended GPT-2 to relation extraction, which is a task of identifying the relationship between concepts appeared in text. They had shown a milestone in predicting larger range of distinct relation types with higher confidence [@alt2019fine]. In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning [@bird2020lstm]. @bird2020lstm employed GPT-2 as part of the solution to generate augmented data for classification. GPT-2 also helped NLP researchers tackling the task of textual paraphrasing. @hegde2020unsupervised used GPT-2 to build an unsupervised paraphrasing model that is not restricted by domain shift within an article, which is a problem faced by the usual supervised models of paraphrasing.

Most of prior language models are trained in a single domain by supervised learning to perform a single task. OpenAI, on the other hands, train the GPT models to perform many tasks [@radford2019language]. GPT-2 is pre-trained on WebText, a comprehensive dataset that is scrapped from millions of webpages and contains total of 40 GB of text [@radford2019language]. WebText contains all kinds of textual data: dialogues, comments, translations, questions and answers...etc. Such diversity is critical for GPT-2 to be a generalized model that can perform multiple tasks with any domain boundary. However, GPT-2 has the capacity to adapt to a context and generates custom output with the "fine-tuning" technique [@ziegler2019fine]. While the pre-training stage is unsupervised and mandatory, fine-tuning is an optional training stage where GPT can learn to perform new tasks in a supervised or reinforced approach such as stylistic continuation and summarization [@ziegler2019fine]. During the pre-training stage, the model is trained to perform text completion task. Give it a prompt to start and it will generate texts that continues what has been started. In the fine-tuning stage, @radford2019language provided additional training datasets for GPT-2 to generate the texts with specific sentiment, or be descriptive, or even summarize the prompt text.

OpenAI has made GPT-2 open-sourced through Github with Python and Tensorflow [@tensorflow2015-whitepaper]. There is also a package in R that wraps the Python code into a R package, created by @keydanaluraschi2019gpt2. There are three choices of models in terms of parameter sizes — "124M", "355M" and "774M". The R package only permits exploratory experiments with the pre-trained tasks; fine-tuning for new tasks is not available thorough R. To proceed with the implementation of GPT-2 in R, theoretically, we will fine-tune the model through Python and create a wrapper code for R. This could possibility be achieved by employing the `reticulate` R package [@reticulateR] that allows operation of Python code in R. The practicality of this approach is to be further investigated.

GPT-3 has over 175 billions of parameters [@brown2020language]. In the past few years, the growth of parameters of transformer language models expands rapidly. The first generation GPT has 110 million parameters. Released in 2018, GPT-2 has 1.6 billion parameters, and within two years, GPT-3 has grown the size to 175 billion, which is over 100 times than its predecessor. The more parameters there are the better the model can generalize to new tasks; therefore, the increasing amount of parameters accelerates the improvements in text synthesis and downstream NLP tasks [@brown2020language]. GPT-3 can learn NLP tasks with just a handful of example; OpenAI refers this ability as few-shot learning, or in-context learning [@brown2020language]. As a result, we only need to provide GPT-3 a few demonstrations for it to learn a new NLP task. For example, OpenAI team trained GPT-3 to generate "news articles". The prompt input is a plausible first sentence of a news story written by human, and the output is an entire news article generated by the model itself. Without the demonstrations, GPT-3 tends to treat the prompts as "tweets" and generate texts that serve as responses or follow-up tweets. The team fed a few previous news articles in the model’s context as demonstrations, and the model learned to produce the outputs that adapt to the style and length of news stories [@brown2020language]. This few-shot learning ability indicates that the model can learn like human do and would democratize the use of language models to broader fields.

@@

BERT stands for Bidirectional Encoder Representations from Transformers [@devlin2018bert]. It was released by researchers at Google AI Language in 2018. The key innovation of BERT is its application of bidirectional training of the Transformer model, which results in a deeper sense of language context and flow. What is bidirectional training? In the context of English language, a sentence is constructed from left to right. The context on the left is used to determine the next word on the right. In this sense, GPT models are unidirectional since the probability of the next word is dependent only on the past. It is also possible to reverse the direction and model the sentence from right to left, and BERT incorporates both directions of left-to-right and right-to-left to model bidirectional contexts.

BERT is pre-trained by using two unsupervised tasks. The first task is Masked Language Model and and second task is Next Sentence Prediction. The bidirectional approach of BERT is alleviated by the "Masked Language Model" [@devlin2018bert]. The Masked Language Model masks certain words in the input corpus which forces the model to predict the original words by "filling in the blank". This training process blends the representation from the left and right context and therefore increases the model's context sensitivity [@nozza2020mask]. In the Next Sentence Prediction process, the model receives two sentences in sequence as input. 50% of the sentences are paired in the right sequence and the ones from the other half have sentences that are paired randomly [@devlin2018bert]. The goal of this process is to have BERT evaluate if the second sentence is the next sentence of the first sentence in the original text and sharpen its capacity to predict the likelihood of the next sentence.

BERT was pre-trained on BooksCorpus dataset with 800 million words and English Wikipedia with 2,500 million words [@devlin2018bert]. Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available with one additional output layer and a task-specific input [@devlin2018bert]. Google AI researchers have fine-tuned BERT for 11 NLP tasks including question-answering and common-sense inference.

BERT implementation in R is also possible through Keras. @kerasbert created a python module called `keras-bert` and the module can be called by R through `reticulate` package [@reticulateR], which enables the operation of Python modules in R. To start, we install the keras-bert model in Python on the machine and load BERT model into R through `reticulate`. Once `keras-bert` is loaded, we can use it to fine-tune BERT on a task-specific training dataset. Fine-tuning is started by tokenizing the training data. The output of the tokenization is then fed to BERT as the input. The next step is to define the model by configuring its parameters such as batch size, number of epochs and the learning rate. Finally, the model is compiled and we can begin the fine-tuning.
